Formalization of Abstract Rewriting

### 2025-11-05 Relations/Paper/Introduction.tex
related work and contributions merged

\subsection{Contributions}
Our contributions are as follows.
\begin{enumerate} \label{list:ourContributions}
  \item An original formalization of elementary ARS theory as presented in
  \cite{Terese}, including relevant background; 
  % Chapter 1 of
  \item An ontology of termination and confluence properties, and a detailed analysis of logical relationships between them;
  \item Definitions of new ARS concepts that refine our understanding of these relationships;
  \item Marginal improvements to classical confluence and termination criteria;
  \item An examination of several distinct notions of well-foundedness
  in the constructive setting.
\end{enumerate}

\subsection{Related works}
% We should cover the following in this section:
% \begin{itemize}
%   \item The IsaFoR/CeTA tool for certification of termination/confluence proofs provided by some automated tool (http://cl-informatik.uibk.ac.at/software/ceta/#introduction). Specifically this paper (https://www.isa-afp.org/entries/Abstract-Rewriting.html) deals with the formalization of strong normalization and Neman's lemma as part of the CeTA project. They use Isabelle for their formalization. 
%   \item This termination portal which gives an overview of various termination tools: https://termination-portal.org/wiki/Termination_Portal. Specifically here is a list of different termination tools: https://termination-portal.org/wiki/Category:Tools . We should talk about what distinguishes our project from a termination tool. 
%   \item Similarly this project is a collection of formalizations of theorems related to term rewrite systems in the language of the proof assistant PVS. http://trs.cic.unb.br/ (here is the formalization of Newman's lemma if you're interested! https://github.com/nasa/pvslib/blob/master/TRS/newman_yokouchi.pvs). 
%   \item Can nention that the agda standard library has a rewriting formalization too based on TeReSe
% \end{itemize}
Several formalizations of the concepts of elementary ARS theory exist, including projects in Isabelle \cite{isabelleFormalization}, PVS \cite{trs}, and Coq \cite{coqARS}, as well as the Agda standard library \cite{AgdaLibraryRewriting}. Our work distinguishes itself through contributions 2-5 above: the detailed ontology of termination and confluence properties, new ARS concepts, refinements to classical criteria, and  examining multiple notions of well-foundedness.

>>>>

\subsection{Contributions and related works}
Several formalizations of the concepts of elementary ARS theory exist, including projects in Isabelle \cite{isabelleFormalization}, PVS \cite{trs}, and Rocq \cite{coqARS} as well as the Agda standard library \cite{AgdaLibraryRewriting}.
The first three are quite mature large-scale projects for certifying termination of
general rewrite systems.  Our work is more comparable in scale to the rewriting development in the
Agda library, and is meant to serve as a starting point for a new library of formalized
programming language theory.  The specific contributions we provide beyond the works cited above
are as follows.
% Our work distinguishes itself through contributions 2-5 above: the detailed ontology of termination and confluence properties, new ARS concepts, refinements to classical criteria, and  examining multiple notions of well-foundedness.
% Our contributions are as follows.
\begin{itemize} \label{list:ourContributions}
  \item An original formalization of elementary ARS theory as presented in Terese
  \cite{Terese}, including relevant background about relations and logic;
  % Chapter 1 of
  \item An ontology of termination and confluence properties, and a detailed analysis of logical relationships between them;
  \item Definitions of new ARS concepts that refine our understanding of these relationships;
  \item Marginal improvements to classical confluence and termination criteria;
  \item An examination of several distinct notions of well-foundedness
  in the constructive setting.
\end{itemize}

% \subsection{Related works}
% We should cover the following in this section:
% \begin{itemize}
%   \item The IsaFoR/CeTA tool for certification of termination/confluence proofs provided by some automated tool (http://cl-informatik.uibk.ac.at/software/ceta/#introduction). Specifically this paper (https://www.isa-afp.org/entries/Abstract-Rewriting.html) deals with the formalization of strong normalization and Neman's lemma as part of the CeTA project. They use Isabelle for their formalization.
%   \item This termination portal which gives an overview of various termination tools: https://termination-portal.org/wiki/Termination_Portal. Specifically here is a list of different termination tools: https://termination-portal.org/wiki/Category:Tools . We should talk about what distinguishes our project from a termination tool.
%   \item Similarly this project is a collection of formalizations of theorems related to term rewrite systems in the language of the proof assistant PVS. http://trs.cic.unb.br/ (here is the formalization of Newman's lemma if you're interested! https://github.com/nasa/pvslib/blob/master/TRS/newman_yokouchi.pvs).
%   \item Can nention that the agda standard library has a rewriting formalization too based on TeReSe
% \end{itemize}
